---
title: "Week 5 Applications"
date: "2025-02-13"
output: 
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 10
    fig_height: 4
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages
library(tidyverse)
library(rvest)
library(polite)
library(httr)
library(jsonlite)
```

# Web scrapping
## Demographics in Singapore

In the first practice,we shall visit the page https://en.wikipedia.org/wiki/Demographics_of_Singapore, which contain multiple tables for demographics in Singapore.

We will obtain the tables on

+ Gender composition of resident population, and

+ household income from work.

```{r}
url <- "https://en.wikipedia.org/wiki/Demographics_of_Singapore"
url_use <- bow(url)
results <- scrape(url_use) %>% 
  html_elements("table.wikitable") %>% html_table()
gender_compo <- results[[5]]
household_income <- results[[21]]
```

```{r}
gender_compo
```

```{r}
household_income
```

## Gross monthly income in Singapore

Next, visit the page

https://stats.mom.gov.sg/Pages/Income-Summary-Table.aspx

And obtain the first table on the median gross monthly income from employment.

```{r}
url <- "https://stats.mom.gov.sg/Pages/Income-Summary-Table.aspx"
url_use <- bow(url)
selector <- "#iMAS_SP_Summ"
tables <- scrape(url_use) %>% html_elements(selector) %>% html_table() # list of 6
df <- tables[[1]]
df
```

# Working with APIs
## Real median household income

As API calls to FRED requires an API key, we need to sign up for an API key first.

+ In the code below, my API key is retrieved in `Sys.getenv("FRED_KEY")`. This is one way to keep the API key secret. You can find out more about it [here](https://daattali.gitbooks.io/stat545-ubc-github-io/content/bit003_api-key-env-var.html).

```{r}
resource_url <- "https://api.stlouisfed.org/fred/series/observations"
query_params = list(api_key = Sys.getenv("FRED_KEY"),
                    series_id = "MEHOINUSA672N",
                    file_type = "json")
# Request data from the server
res = GET(resource_url, query = query_params)

# Parse JSON
res_list <- content(res, type = "text") %>% fromJSON(flatten = TRUE)

# Convert data into tibble
income <- as_tibble(res_list$observations)
head(income)
```

```{r}
income %>%
  mutate(year = year(date), value = as.numeric(value)) %>%
  filter(year >= 1990) %>%
  ggplot(aes(x = year, y = value)) +
  geom_line(lwd = 2, color = "indianred4", alpha = 0.9) +
  theme_minimal() +
  labs(x = "", y = "", title = "Real Median Household Income in the US, 1990 to 2022",
       caption = "Source: U.S. Census Bureau.")
```

## Realtime Carpark availability

To query data from the LTA DataMall, we also need to request an API key.

+ In the code below, my API key is retrieved in `Sys.getenv("LTA_KEY")`.

+ Read the LTA's **documentation** for their APIs and understand how to set up the `GET()` requests.

```{r}
# Construct the resource URL
resource_url <- "http://datamall2.mytransport.sg/ltaodataservice/CarParkAvailabilityv2"
# Make the GET() request
res <- GET(resource_url, 
           add_headers(AccountKey = Sys.getenv("LTA_KEY"),
                       accept = "application/json"))
# Parse the returned data
res_content <- content(res, as = "text")
res_list <- fromJSON(res_content, flatten = TRUE)
# Extract information from sub-list
df_carpark <- as_tibble(res_list$value)
head(df_carpark)
```

### The `ltaer` package

there's also an API wrapper `R` package, `ltaer`, that helps us obtain data from the LTA DataMall.

+ Read the full documentation of the package [here](https://shaunkhoo.github.io/ltaer/index.html).

+ To obtain data via this API wrapper, we still need to supply our own API KEY.

+ The data set on carpark availability can be retrieved with:

```{r}
# Retrieve data on carpark availability
carpark_avail <- ltaer::getCarparkAvail(Sys.getenv("LTA_KEY"))
head(carpark_avail)
```

### Car park availability: Visualization

Lastly, let's try visualize the real-time car park availability. We will need an additional package, [leaflet](https://rstudio.github.io/leaflet/).

```{r eval = FALSE}
library(leaflet)
df_carpark <- carpark_avail %>%
  separate(Location, into = c("lat", "long"), sep = " ", convert = TRUE)

# Visualization
leaflet(df_carpark) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~long, lat = ~lat,
                   popup = ~paste0("Available slots: ", AvailableLots),
                   radius = ~AvailableLots/100, stroke = FALSE, fillOpacity = 0.5)
```

